---
title: "Analyzing Tidy Tuesday Star Wars Survey Dataset"
subtitle: ""
output:
  github_document
always_allow_html: true
editor_options:
  chunk_output_type: console
---

## Setup

```{r libs, message=FALSE, warning=FALSE}

library(tidyverse)
library(here)
library(kableExtra)

```

```{r theme_chunk_opts}

theme_set(theme_light())

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE,
                      fig.width = 10,
                      fig.height = 7)

```

=

```{r star_wars_survey_raw}

star_wars_survey_raw <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-05-14/week7_starwars.csv")

```

```{r star_wars_survey}

star_wars_survey <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2018/2018-05-14/week7_starwars.csv") %>%
  select(-32) %>% 
  rename(seen_any_star_wars = `Have you seen any of the 6 films in the Star Wars franchise?`,
         fan_of_star_wars = `Do you consider yourself to be a fan of the Star Wars film franchise?`,
         seen_ep_I = `Which of the following Star Wars films have you seen? Please select all that apply.`,
         seen_ep_II = X5,
         seen_ep_III = X6,
         seen_ep_IV = X7,
         seen_ep_V = X8,
         seen_ep_VI = X9,
         ep_I_rank = `Please rank the Star Wars films in order of preference with 1 being your favorite film in the franchise and 6 being your least favorite film.`,
         ep_II_rank = X11,
         ep_III_rank = X12,
         ep_IV_rank = X13,
         ep_V_rank = X14,
         ep_VI_rank = X15,
         han_fav = `Please state whether you view the following characters favorably, unfavorably, or are unfamiliar with him/her.`,
         luke_fav = X17,
         leia_fav = X18,
         anakin_fav = X19,
         obi_won_fav = X20,
         palpy_fav = X21,
         vader_fav = X22,
         lando_fav = X23,
         boba_fett_fav = X24,
         c3po_fav = X25,
         r2d2_fav = X26,
         jar_jar_fav = X27,
         padme_fav = X28,
         yoda_fav = X29,
         which_char_shot_first = `Which character shot first?`,
         familiar_with_expanded = `Are you familiar with the Expanded Universe?`,
         #fan_of_expanded = `Do you consider yourself to be a fan of the Expanded Universe?\u008c\xe6`,
         star_trek_fan = `Do you consider yourself to be a fan of the Star Trek franchise?`,
         gender = Gender,
         age_group = Age,
         income = `Household Income`,
         education = Education,
         region = `Location (Census Region)`) %>% 
  slice(2:nrow(.)) %>% 
  mutate(age_group = fct_relevel(age_group, c(NA, "18-29", "30-44", "45-60", "> 60"))) %>% 
  mutate_at(.vars = vars(seen_ep_I:seen_ep_VI), .funs = ~ifelse(is.na(.), 0, 1))

```

## Character-specific questions

First, I create a dataset that contains character-specific questions and is formatted so that each row represents a single respondent's answers on a single character. This way favorability information can easily be plotted with ggplot.

```{r character_questions}

character_questions <- star_wars_survey %>%  
  select(-contains("ep")) %>% 
  pivot_longer(cols = contains("fav"),
               names_to = "character", 
               values_to = "favorability") %>%
  drop_na(favorability) %>% 
  mutate(favorability = ifelse(favorability == "Neither favorably nor unfavorably (neutral)", "Neutral", favorability)) %>% 
  mutate(character = str_remove(character, "_fav"),
         favorability = fct_relevel(favorability, c("Unfamiliar (N/A)",
                                                    "Very unfavorably", 
                                                    "Somewhat unfavorably", 
                                                    "Neutral", 
                                                    "Somewhat favorably",
                                                    "Very favorably")),
         
         # create a favorability score so facets can be ordered from most to least favorable and so quantitative favorability differences can be calculated
         favorability_score = case_when(favorability == "Unfamiliar (N/A)" ~ 0,
                                        favorability == "Very unfavorably" ~ -3,
                                        favorability == "Somewhat unfavorably" ~ -2,
                                        favorability == "Neutral" ~ -1,
                                        favorability == "Somewhat favorably" ~ 2,
                                        favorability == "Very favorably" ~ 4,
                                        TRUE ~ -9999))

character_questions %>% 
  head(5) %>% 
  kable() %>% 
  kable_styling()

```

### Who are the most popular characters?

```{r character-fav-overall}

character_questions %>% 
  mutate(character = fct_reorder(character, favorability_score, .desc = TRUE)) %>% 
  ggplot(aes(favorability, fill = favorability)) +
  geom_bar() +
  facet_wrap(~ character) +
  scale_fill_manual(values = c("black", "red", "red4", "lightgrey", "green4", "green")) +
  coord_flip() +
  labs(title = "Favorability of Star Wars Characters",
       x = "Favorability",
       y = "Count") +
  theme(legend.position = "none")

ggsave(here("star-wars-survey", "plots", "character-fav-overall.png"))

```

This is a great demonstration of what vague survey questions can lead to :)

People love Vader but gave him bad "favorability" because he's an asshole, whereas the good guys from the originals all grade out well. The "good guys" from the prequels (Padme and Anakin) have middling reviews as well, probably because so many people hate those movies.

### Which characters see favorability differences between different age groups?

```{r character_fav_by_age_group}

character_fav_by_age_group <- character_questions %>% 
  group_by(character, age_group) %>% 
  summarise(n = n(), fav_score = mean(favorability_score, na.rm = TRUE)) %>% 
  drop_na()

```

```{r character-fav-by-age-group}

character_fav_by_age_group %>%
  ggplot(aes(age_group, fav_score, fill = age_group)) +
  geom_col() +
  facet_wrap(~ character) +
  coord_flip() +
  ylim(-3, 4) +
  labs(title = "Favorability of character by age group",
       x = "Age group",
       y = "Favorability score")

ggsave(here("star-wars-survey", "plots", "character-fav-by-age-group.png"))

```

A few conclusions:

- Younger people feel less strongly favorable towards Original Trilogy characters (C3PO, Han, Luke) and more favorably towards OT villians (Vader, Boba Fett)
- Older people are more ambivalent towards Jar-Jar!

### Do fans familiar with the Expanded Universe feel differently about any characters?

I'll first get a value for each character indicating how much more positively those who are familiar with the Expanded Universe feel about them.

```{r expanded_fav_diff}

expanded_fav_diff <- character_questions %>% 
  group_by(character, familiar_with_expanded) %>% 
  summarise(n = n(), fav_score = mean(favorability_score)) %>% 
  drop_na() %>% 
  pivot_wider(id_cols = "character", names_from = "familiar_with_expanded", values_from = "fav_score") %>% 
  mutate(expanded_fav_diff = Yes - No) %>% 
  select(character, expanded_fav_diff)

```

```{r character-fav-by-exp-univ}

character_questions %>% 
  group_by(character, familiar_with_expanded) %>% 
  summarise(n = n(), fav_score = mean(favorability_score)) %>% 
  ungroup() %>% 
  drop_na() %>%
  inner_join(expanded_fav_diff, by = "character") %>%
  mutate(character = fct_reorder(character, expanded_fav_diff),
         familiar_with_expanded = fct_relevel(familiar_with_expanded, "Yes", "No")) %>% 
  ggplot(aes(character, fav_score, color = familiar_with_expanded)) +
  geom_point(size = 5) +
  coord_flip() +
  ylim(-3, 4) +
  labs(title = "Character favorability for those who are and are not familiar with the Expanded Universe",
       y = "Favorability Score",
       x = "")

ggsave(here("star-wars-survey", "plots", "character-fav-by-exp-univ.png"))

```

I was hoping to find a few characters like Boba Fett and Palpatine, who I know are more fleshed out in the EU, appear at the top! I'm not sure how many EU stories expand on Vader, but this leads me to think there are quite a few.

## Film-specific questions

I again create a subsetted long-form dataset, but this time it contains film-specific questions.

```{r saw_ep_long}

saw_ep_long <- star_wars_survey %>% 
  select(-contains("fav"), -contains("rank")) %>% 
  pivot_longer(cols = seen_ep_I:seen_ep_VI,
               names_to = "film",
               values_to = "saw_ep") %>% 
  mutate(film = str_remove(film, "seen_"))
  

```

```{r ep_rank_long}

ep_rank_long <- star_wars_survey %>% 
  select(-contains("fav"), -contains("seen_ep")) %>% 
  pivot_longer(cols = ep_I_rank:ep_VI_rank,
               names_to = "film",
               values_to = "ep_rank") %>% 
  mutate(film = str_remove(film, "_rank"))

```

```{r film_questions}

film_questions <- inner_join(saw_ep_long, ep_rank_long)

# shows data for 2 respondents (12 rows because 2 * 6 movies = 12 rows for 2 respondents)
film_questions %>% 
  head(12) %>% 
  kable() %>% 
  kable_styling()

```

### What is the distribution of film rank for each movie?

```{r movie-fav-overall}

film_questions %>% 
  filter(saw_ep == 1) %>% 
  mutate(film = fct_relevel(film, c("ep_I", "ep_IV", "ep_II", "ep_V", "ep_III", "ep_VI")),
         ep_rank = fct_relevel(ep_rank, c("6", "5", "4", "3", "2", "1"))) %>%
  drop_na() %>% 
  ggplot(aes(ep_rank, fill = ep_rank)) +
  geom_bar(stat = "count") +
  facet_wrap(~ film, scales = "free_y", ncol = 2) +
  coord_flip() +
  scale_fill_manual(values = c("red4", "red3", "red", "green", "green3", "green4")) +
  labs(title = "Favorability of each movie",
       x = "Episode rank",
       y = "Count") +
  theme(legend.position = "none")

ggsave(here("star-wars-survey", "plots", "movie-fav-overall.png"))

```

Staunch differences between the OT and the PT. Also interesting that 4 and 5 have a similar total of positive/negative votes, but 5 is the overall favorite for waaay more people.

### And what about for different age groups?

```{r movie-fav-by-age}

film_questions %>% 
  filter(saw_ep == 1) %>% 
  mutate(film = fct_relevel(film, c("ep_I", "ep_II", "ep_III", "ep_IV", "ep_V", "ep_VI")),
         ep_rank = fct_relevel(ep_rank, c("6", "5", "4", "3", "2", "1"))) %>%
  drop_na() %>% 
  ggplot(aes(ep_rank, fill = ep_rank)) +
  geom_bar(stat = "count") +
  facet_grid(age_group ~ film, scales = "free_y") +
  coord_flip() +
  scale_fill_manual(values = c("red4", "red3", "red", "green", "green3", "green4")) +
  labs(title = "Favorability of each movie among different age groups",
       x = "Episode rank",
       y = "Count") +
  theme(legend.position = "none")

ggsave(here("star-wars-survey", "plots", "movie-fav-by-age.png"))

```

Not much is surprising about the OT movies here with older groups favoring them slightly more. The Prequel movies are odd though...Episodes 2 and 3 follow a pattern I expected: slightly higher favorability among younger folks but still negative. So my question from this is: Why do 45+ year olds like Episode 1 so much?

## Appendix

### Which movies have been seen by the most people?

```{r}

film_questions %>% 
  mutate(film = fct_reorder(film, saw_ep, .fun = mean, .desc = TRUE)) %>%
  group_by(film, age_group) %>% 
  summarise(pct_seen = sum(saw_ep) / n()) %>% 
  ungroup() %>% 
  drop_na() %>%
  ggplot(aes(film, pct_seen, fill = age_group)) +
  geom_col() +
  coord_flip() +
  labs()

```

Not much interesting to see here besides the overall rank. I'm surprised that more have seen 5 and 6 than 4, though it goes to show how much Star Wars caught on after A New Hope. The Prequel Trilogy, on the other hand, started poorly with Episode 1 and so it's unsurprising that fewer people ended up seeing 2 and 3.



